The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-03-24 VegOzone.tex
Finished captions and spell check.

\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subartic climate}

@@

\begin{abstract}
  \begin{itemize}
  \item ozone damage
  \item present subarctic climate and conditions
    \begin{itemize}
    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
    \item[$\rightarrow$] ozone climatology and spring peak
    \end{itemize}
  \item future climate and conditions
    \begin{itemize}
    \item[$\rightarrow$] increase in heatwaves and dry periodes
    \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
    \item[$\rightarrow$] increase in tropospheric background ozone
    \end{itemize}
  \item present day test case 2018 vs 2019 ozone induced damage at NIBIO research station Svanhovd
  \item findings and results 
  \end{itemize}
\end{abstract}

@@

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
\begin{itemize}
\item ozone in general
\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
\item present subarctic climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
  \item[$\rightarrow$] ozone climatology and spring peak
  \end{itemize}
\item future climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] increase in heatwaves and dry periodes
  \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
  \item[$\rightarrow$] increase in tropospheric background ozone
  \end{itemize}
\end{itemize}

@@

\subsection{Observational data}
Table: EBAS - Station list: ID - name - location - country - operational
\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Timeseries of ozone observations in northern Fennoscadia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19. The hatched areas indicate time periods with unsufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}

\subsection{Model and reanalysis data}
Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
Precipitation/Temperature: CRU biascorrected rea
Ozone uptake / POD: DO3SE

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia. For comparison also data from Prestbakke (southern Norway) is shown. The large spread in case of Svanvik and Janiskoski is due to the lower statistics. All northern Fennoscandic climatologies peak in spring, with peak values in April. In southern Norway, peak values appear only in May. The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
  \label{fig:ozone_climatology_fenoscandic_obs}
\end{figure}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatologies}
  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscadia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanlysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed surface ozone well. The MACC reanalysis is in general too low and displays a erronous seasonal cycle. The CAMS reanalysis matches quite well in southern Fennoscandia but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscadia, though the ozone concentrations in the north are slightly to low.}
  \label{fig:ozone_climatologies}
\end{figure}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
  \caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
  \label{fig:ozone_climatology_fenoscandic_obs_test}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
  \label{fig:plot_temperature_anomalies_svanvik}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
  \label{fig:plot_precipitation_anomalies_svanvik}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\section{DO3SE modeling}
\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
  \label{fig:ozone_obs_2018_07}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}
Model results for natural and semi-natural vegetation (birch, pine, clover)
\begin{itemize}
\item climatological ozone impact
\item 2018
\item 2019
\end{itemize}

@@

\appendixfigures  %% needs to be added in front of appendix figures
\begin{figure}[t]
  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
  \label{fig:greening_season_change_Svanvik}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{test_fennoscandia}
  \label{fig:sunlight_fennoscandia}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{density_distribution}
  \label{fig:density_distribution}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
  \label{fig:ozone_obs_2018_rrea}
\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
\end{figure}

>>>>

\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}

@@

\begin{abstract}
  \begin{itemize}
  \item ozone damage
  \item present subarctic climate and conditions
    \begin{itemize}
    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
    \item[$\rightarrow$] ozone climatology and spring peak
    \end{itemize}
  \item future climate and conditions
    \begin{itemize}
    \item[$\rightarrow$] increase in heatwaves and dry periods
    \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
    \item[$\rightarrow$] increase in tropospheric background ozone
    \end{itemize}
  \item present day test case 2018 vs 2019 ozone induced damage at NIBIO research station Svanhovd
  \item findings and results 
  \end{itemize}
\end{abstract}

@@

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
\begin{itemize}
\item ozone in general
\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
\item present subarctic climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
  \item[$\rightarrow$] ozone climatology and spring peak
  \end{itemize}
\item future climate and conditions
  \begin{itemize}
  \item[$\rightarrow$] increase in heatwaves and dry periods
  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
  \item[$\rightarrow$] increase in tropospheric background ozone
  \end{itemize}
\end{itemize}

@@

\subsection{Observational data}
Table: EBAS - Station list: ID - name - location - country - operational
\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Time series of ozone observations in northern Fennoscandia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19. The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}

\subsection{Model and reanalysis data}
Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
Precipitation/Temperature: CRU bias corrected rea
Ozone uptake / POD: DO3SE

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia. For comparison also data from Prestebakke (southern Norway) is shown. The large spread in case of Svanvik and Janiskoski is due to the lower statistics. All northern Fennoscandic climatologies peak in spring, with peak values in April. In southern Norway, peak values appear only in May. The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
  \label{fig:ozone_climatology_fenoscandic_obs}
\end{figure}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatologies}
  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean). As representative for the sites, the nearest neighboring grid point has been chosen. The ECMWF products do not reproduce observed surface ozone well. The MACC reanalysis is in general too low and displays a erroneous seasonal cycle. The CAMS reanalysis matches quite well in southern Fennoscandia but does not reproduce the seasonality in northern Fennoscandia. The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscandia, though the ozone concentrations in the north are slightly to low.}
  \label{fig:ozone_climatologies}
\end{figure}

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas; (c) Prestebakke.}
  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level. (a) Esrange; (b) Pallas; (c) Prestebakke.}
  \label{fig:ozone_climatology_fenoscandic_obs_test}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively. \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
  \caption{Time series over the past 6 years of (a) temperature observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of temperature. The summer of 2018 has been significantly warmer on the $2-4\,\sigma$ level. The summer of 2019 was fairly average but on the colder side.}
  \label{fig:plot_temperature_anomalies_svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
  \caption{Time series over the past 6 years of (a) precipitation observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of precipitation.}
  \label{fig:plot_precipitation_anomalies_svanvik}
\end{figure}

\section{DO3SE modeling}

Model results for natural and semi-natural vegetation (birch, pine, clover)
\begin{itemize}
\item climatological ozone impact
\item 2018
\item 2019
\end{itemize}

@@

\appendixfigures  %% needs to be added in front of appendix figures
\begin{figure}[t]
  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
  \label{fig:greening_season_change_Svanvik}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{test_fennoscandia}
  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia. Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July. For comparison with southern Fennoscandia, Prestebakke is shown.}
  \label{fig:sunlight_fennoscandia}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{density_distribution}
  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia. Jergul/Karasjok is well correlated with Esrange, and Pallas. Svanvik displays highest correlation with Pallas. (a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
  \label{fig:density_distribution}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean. (a) Time series; (b) Student's t-test.}
  \label{fig:ozone_obs_2018_rrea}
\end{figure}
